---
title: "Class 11: Structural Bioinformatics (Pt1)"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Revist the PDB

> Q1: Download a CSV file from the PDB site (accessible from “Analyze” -> “PDB Statistics” > “by Experimental Method and Molecular Type”. Move this CSV file into your RStudio project and determine the percentage of structures solved by X-Ray and Electron Microscopy. Also can you determine what proportion of structures are protein?

Downloaded a CSV file from: http://www.rcsb.org/stats/summary on 2019-05-08 

```{r}
db <- read.csv("Data Export Summary.csv", row.names = 1)
head(db)
```

```{r}
## Total number of structures in the PDB database
sum(db$Total)
```

```{r}
# Precent of each exper method
(db$Total/sum(db$Total)) * 100
```

What precent are protein?

```{r}
(sum(db$Proteins) / sum(db$Total)) * 100
```



## Section 3. Using Bio3D

Let's read the PDB file we downloaded into R to see how things live in this place

```{r}
library(bio3d)
```

```{r}
pdb <- read.pdb("1hsg.pdb")
pdb
```

```{r}
aa321(pdb$seqres)
```


```{r}
head(pdb$atom)
```

### Atom selection examples

Lets select residue 10
```{r}
inds <- atom.select(pdb, resno = 10)
```


```{r}
pdb$atom[inds$atom, ]
```


```{r}
atom.select(pdb, resno=10, value = TRUE)
```

### Make a protein only and ligand only PDB file

Here we need to select our protein with `atom.select()` then use `write.pdb()`


```{r}
prot <- atom.select(pdb, "protein", value=TRUE)
prot
```

```{r}
write.pdb(prot, file="1hsg_protein.pdb")
```

```{r}
lig <- atom.select(pdb, "ligand", value=TRUE)
lig
```

```{r}
write.pdb(lig, file="1hsg_ligand.pdb")
```


```{r}
library(bio3d.view)

view(lig)
```

```{r}
view(pdb, "overview", col="sse")

```

## Section 6. Working with multiple PDBs

```{r}
# Download some example PDB files
ids <- c("1TND_B","1AGR_A","1TAG_A","1GG2_A","1KJY_A","4G5Q_A") 

files <- get.pdb(ids, split = TRUE)
```

Align and superpose these molecules
```{r}
pdbs <- pdbaln(files, fit=TRUE)
```

View them all supperposed
```{r}
view(pdbs)
```

Perform PCA on the molecule set
```{r}
pc.xray <- pca(pdbs)
```

```{r}
# PLot of PCA results
plot(pc.xray)
```

```{r}
 # Visualize first principal component
pc1 <- mktrj(pc.xray, pc=1, file="pc_1.pdb")
```

```{r}
view(pc1)
```


