---
title: 'Class 9: Unsupervised Learning Analysis of Human Breast Cancer Cells'
author: "Barry Grant"
date: "5/1/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Preparing the data

I downloaded the CSV format data file from the class website and placed it in my project directory

```{r}
wisc.df <- read.csv("WisconsinCancer.csv")
head(wisc.df)
```

> Q3. How many of the observations have a malignant diagnosis?

```{r}
table(wisc.df$diagnosis)
```

> Q4. How many variables/features in the data are suffixed with _mean?

```{r}
length( grep("_mean", colnames(wisc.df)) )
```

Select cols 3:32 and store as a matrix

```{r}
wisc.data <- as.matrix(wisc.df[,3:32])
# Set the row names of wisc.data
row.names(wisc.data) <- wisc.df$id

```


```{r}
# Create diagnosis vector for later 
diagnosis <- wisc.df$diagnosis
```



## Principal Component Analysis (PCA)

Do I need to scale the data before PCA? Check the col means and standard deviations...

```{r}
round( apply(wisc.data, 2, mean), 1 )
```

```{r}
round( apply(wisc.data, 2, sd), 1 )
```


Looks like I need to scale this data before PCA or just set the `scale=TRUE` argument to `prcomp()`

```{r}
wisc.pr <- prcomp(wisc.data, scale=TRUE)
summary(wisc.pr)
```


```{r}
# Scatter plot observations by components 1 and 2
plot( wisc.pr$x[,1], wisc.pr$x[,2] , col = diagnosis, 
     xlab = "PC1", ylab = "PC2")
```


```{r}
# Scatter plot observations by components 2 and 3
plot( wisc.pr$x[,2], wisc.pr$x[,3] , col = diagnosis, 
     xlab = "PC2", ylab = "PC3")
```


> Q13. For the first principal component, and using two significant figures , what is the component of the loading vector (i.e. wisc.pr$rotation[,1]) for the feature radius_mean?


```{r}
round( wisc.pr$rotation["radius_mean",1], 2)
```

> Q14. For the first principal component, and using two significant figures, what is the component of the loading vector (i.e. wisc.pr$rotation[,1]) for the feature smoothness_se?

```{r}
round( wisc.pr$rotation["smoothness_se",1], 2)
```

> Q15. Which original variable contributes most to PC1?

```{r}
sort(abs(wisc.pr$rotation[,1]))
```



## Hierarchical clustering

First we need to scale or data as we saw previously, then distance matrix then hierarchical clustering...

```{r}
wisc.hclust <- hclust( dist( scale(wisc.data) ) )
```


```{r}
plot(wisc.hclust)
abline(h=19, col="blue")
```

```{r}
wisc.hclust.clusters <- cutree(wisc.hclust, k=4)
table(wisc.hclust.clusters)
```

```{r}
table(wisc.hclust.clusters, diagnosis)
```

## Section 5. Combining methods (PCA + hclust)

I am going to start with the PCs that capture 90% of the origional varance in the dataset (i.e. PC1 to PC7)

```{r}
wisc.pr.hclust <- hclust( dist( wisc.pr$x[,1:7] ), method="ward.D2" )
```

```{r}
plot(wisc.pr.hclust)
abline(h=70, col="red", lty=2)
```


```{r}
grps <- cutree(wisc.pr.hclust, k=2)
table(grps)
```

```{r}
table(grps, diagnosis)
```

```{r}
plot(wisc.pr$x[,1], wisc.pr$x[,2], col=grps)
```



