---
title: "Genome Play"
author: "Barry Grant"
date: "4/29/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## The BSGenome package

To install this package, start R (version "3.5") and enter:

```{r}
#BiocManager::install("BSgenome")
```

See the man page for the available.genomes function (?available.genomes) for more information about how to get the list of all the BSgenome data packages currently available in your version of Biocon- ductor

```{r}
library(BSgenome)

available.genomes()
```

Then install one

```{r}
#BiocManager::install("BSgenome.Scerevisiae.UCSC.sacCer3")
```

```{r}
# load the package and store data into yeast
library(BSgenome.Scerevisiae.UCSC.sacCer3)
yeast <- BSgenome.Scerevisiae.UCSC.sacCer3
yeast
```

```{r}
yeast$chrI
```

```{r}
length(yeast)
```

```{r}
length(yeast$chrI)
```

```{r}
# Sequence lengths
seqlengths(yeast)
```

```{r}
# DNA single string
dna_seq <- DNAString("ATGATCTCGTAA")
dna_seq
```

```{r}
# Transcription DNA to RNA string
rna_seq <- RNAString(dna_seq)
rna_seq
```


Three RNA bases form one AA: AUG = M, AUC = I, UCG = S, UAA = *

```{r}
# Translation RNA to AA
aa_seq <- translate(rna_seq)
aa_seq
```

## The Zika virus

It is tiny for a genome
