---
title: "Class 12: Structural Bioinformatics (Pt2)"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Obtain and prepare our structure for 'docking'

Here we download the HIV-Pr structure from the PDB database and extract out the protein and ligand portions to seperate new PDB format files.

```{r}
library(bio3d)

file <- get.pdb("1hsg")
pdb <- read.pdb(file)
pdb
```

```{r}
# Select protein only
prot <- atom.select(pdb, "protein", value=TRUE)
prot
write.pdb(prot, file="1hsg_protein.pdb")
```

```{r}
# Select ligand only
lig <- atom.select(pdb, "ligand", value=TRUE)
lig
write.pdb(lig, file = "1hsg_ligand.pdb")
```

## Add Hydrogens and charges in ADT

We opened our protein only PDB file in AutoDocTools and added hydrogens and atom-types needed for docking calculations.

## Run docking

We will use AutoDoc Vina here at the UNIX command line!
The command I used was:

`Downloads/autodock_vina_1_1_2_mac/bin/vina --config config.txt --log log.txt`

## Process results back in R

```{r}
res <- read.pdb("all.pdbqt", multi=TRUE)
res
write.pdb(res, file="results.pdb")
```

Compare our docking poses to the MERK drug structure. 

```{r}
ori <- read.pdb("ligand.pdbqt")
rmsd(ori, res)
```

## Normal Mode Analysis example

NMA

```{r}
pdb <- read.pdb("1hel") 
modes <- nma( pdb )
m7 <- mktrj(modes, mode=7, file="mode_7.pdb")
```


```{r}
plot(modes)
```

