---
title: "Class 7: R functions and packages"
author: "Barry Grant"
date: "4/24/2019"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## More on function writting

First we will revisit our function from last day

```{r}
source("http://tinyurl.com/rescale-R")
```

Test the **rescale()** function

```{r}
rescale( 1:10 )
```

```{r, eval=FALSE}
rescale( c(1:10, "string") )
```

```{r}
x <- c(1:10, "string")
!is.numeric(x)
```

```{r, eval=FALSE}
rescale2( x )
```

## Function practice

Write a function to identify NA elements in two vectors

Start with a simple example input where I know what the answer should be.

```{r}
x <- c( 1, 2, NA, 3, NA)
y <- c(NA, 3, NA, 3,  4)
```

```{r}
is.na(x)
```

```{r}
is.na(y)
```

I am looking for the positions where it is TRUE in both vectors...

```{r}
is.na(x) & is.na(y)
```

Take the sum to find how many

```{r}
sum( is.na(x) & is.na(y) )
```

This is my working snippet of code that I can use as the body of my first function!

```{r}
both_na <- function(x,y) {
  sum( is.na(x) & is.na(y) )
}
```


```{r}
both_na(x, y)
```

```{r}
both_na(c(NA, NA, NA), c(NA, NA, 1))
```

```{r}
both_na(c(NA, NA, NA), c(1, NA, NA))
```

```{r}
both_na(c(NA, NA, NA), c(1, NA, NA, NA))
```

```{r}
both_na(c(NA, NA, NA), c(1, NA, NA, NA, NA, NA))
```

Check the length of our inputs are equal

```{r}
x <- c(NA, NA, NA)
y <- c(1, NA, NA, NA, NA, NA)
length(x) != length(y)
```


```{r}
3 != 2

```

Try the both_na3() function with extra features

```{r}
x <- c( 1, 2, NA, 3, NA)
y <- c(NA, 3, NA, 3,  4)

both_na3(x, y)
```

```{r}
which(c(F,T,T,F,F,T))
```


## Example grade function

```{r}
x <- c(100, 100, 100, 100, 100, 100, 100, 90)


(sum(x) - min(x)) / (length(x)-1)
```

I got my snippet now to turn it into my first version of the grade() function


```{r}
grade <- function(x) {
  (sum(x, na.rm=TRUE) - min(x, na.rm=TRUE)) / (length(x)-1)
}
```

```{r}
url <- "https://tinyurl.com/gradeinput"

students <- read.csv(url, row.names = 1)
head(students)
```

```{r}
grade(students[5,])
```

```{r}
ans <- apply(students, 1, grade)
```

```{r}
sort(ans, decreasing = TRUE)
```


```{r}
grade(x)
```

```{r}
student2 <- c(100, NA, 90, 90, 90, 90, 97, 80)
grade(student2)
```


## One last function example

Find the intersection of two sets

```{r}
x <- df1$IDs
y <- df2$IDs

intersect(x, y)
```

```{r}
x
y

x %in% y
x[x %in% y]
```

```{r}
y %in% x
```

```{r}
gene_intersect <- function(x, y) {
  cbind( x[x %in% y],
         y[y %in% x] )
}
```

```{r}
merge(df1, df2, by="IDs")
```


```{r}
#install.packages("BiocManager")
```










